\subsection{Compact Constant Folding}
\label{sec:design:constants}

Software Foundations's \emph{Programming Language Foundations} opens with a constant-folding transformation exercise~\cite[Chapter~1]{Pierce:SF2}.
%
Starting from a small language of expressions containing natural numbers, variables, addition, subtraction, and multiplication we are to deploy the semiring properties to simplify expressions.
%
The definition of the simplifying traversal contains much duplicated code due to the way the source language is structured:
%
all the binary operations are separate constructors, whose subterms need to be structurally simplified before we can decide whether a rule applies.
%
The correction proof has just as much duplication because it needs to follow the structure of the call graph of the function it wants to see reduce.
%
The only saving grace here is that Coq's tactics language lets users write scripts that apply to many similar goals thus avoiding duplication in the source file.

In \Velo{}, we structured our core language's representation in such a way that this duplication is never needed.
%
All builtin operators (from primitive operations on builtin types to function application itself) are represented using a single \IdrisData{Call} constructor which takes an operation and a type-indexed list of subterms.

\begin{Verbatim}
data Term : (metas : List Meta) -> (ctxt : List Ty) -> Ty -> Type where
  -- ...
  Call : \{tys : _\} -> (operator : Prim                  tys  ty)
                   -> (operans  : All (Term metas ctxt) tys)
                               -> Term      metas ctxt       ty
\end{Verbatim}

The primitives include:
%
\IdrisData{Zero}---which takes no argument and returns a term of type \IdrisData{TyNat};
%
\IdrisData{And}---which takes two arguments of type \IdrisData{TyBool} and return a term
of type \IdrisData{TyBool};
%
and
%
\IdrisData{App}---which takes a function and an argument that corresponds to the type of the function's domain and returns a term that is the type of the function's co-domain.

\begin{Verbatim}
data Prim : (args : List Ty) -> (type : Ty) -> Type
  where
    Zero : Prim []                    TyNat
    And  : Prim [TyBool, TyBool]      TyBool
    App  : Prim [TyFunc dom cod, dom] cod
    -- ...
\end{Verbatim}

Using \IdrisType{Prim} structural operations can now be implemented by handling recursive calls on the subterms of \IdrisData{Call} nodes uniformly before dispatching on the operator to see whether additional simplifications can be deployed.

Similarly, all of the duplication in the correction proofs is factored out in a single place where the induction hypotheses can be invoked.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:
