%TODO: comment on cost of thinning as inductive values?
%      cite the preprint solving this & whose diagram this is?
\input{ast}

Now that our core language is well scoped by construction, our compiler passes
must also be shown to be scope-preserving.
%
This is not a new requirement, merely it makes concrete a constraint that
used to be enforced informally.
More importantly we show, with our compiler passes, that model-to-same-model transformation of our \ac{edsl} is possible, and that the infrastructure required is not bespoke to \Velo{}.

The purpose of \ac{cse} is to identify subterms that appear multiple times in the syntax tree, and to abstract over them to avoid needless recomputations at runtime.
%
In the following example for instance, we would like to let-bind $t$ before
the application node (denoted \$) so that $t$ may be shared by both subtrees.

\begin{center}
  \cseexamplegraph{}
\end{center}

One of the challenges of \ac{cse}, as exemplified above, is that the term of interest
may be burried deep inside separate contexts.
%
In our intrinsically scoped representation, $t$ in scope $\Gamma, x : \sigma$
is potentially not actually syntactically equal to a copy living in $\Gamma, a : \tau, b : \nu$.
%
Indeed a variable $v$ bound in $\Gamma$ will, for instance, be represented by
the \DeBruijn{} index ($1+v$) in $\Gamma, x : \sigma$
but by the index ($2+v$) in $\Gamma, a :  \tau, b : \nu$.

If only terms were indexed by their exact \emph{support}!
We would not care about additional yet irrelevant variables that happen to be in scope.
%
The principled solution here is to switch to a different representation for
the purpose of \ac{cse}.
The co-\DeBruijn{} representation~\cite{DBLP:journals/corr/abs-1807-04085} provides exactly this guarantee.

%\subsection{Co-\DeBruijn{} representation}

In the co-\DeBruijn{} representation, every term is precisely indexed by its
exact support.
%
That is to say that every subterm explicitly throws away the bound variables
that are not mentioned in it.
%
By the time we reach a variable node, a single bound variable remains in scope:
precisely the one being referred to.

If we think of thinnings as sequences of 0/1 bits stating whether a variable
is kept or dropped, and admitting that such sequences can be represented as
list of either $\bullet$ (1) or $\circ$ (0), the $S$ combinator
($\lambda g. \lambda f. \lambda x. g x (f x)$) is represented as follows in
co-\DeBruijn{} notation (diagram taken from~\cite{MANUAL:draft/Allais22}).

\begin{center}
  \codebruijnexamplegraph{}
\end{center}

The first three $\lambda$ abstractions only use $\bullet$ in their thinnings
because all of $g$, $f$, and $x$ do appear in the body of the combinator.
%
The first application node then splits the context into two: the first subterm
($g x$) drops $f$ while the second ($f x$) gets rid of $g$.
%
Further application nodes select the one variable still in scope for each
leaf subterm: $g$, $x$, $f$, and $x$.

Using a co-\DeBruijn{} representation, we can easily identify shared subterms:
they need to not be mentioning any of the most local variables and be
syntactically equal.


Our pass succesfully transforms the program on the left-hand side to the
one on the right-hand side where the repeated expressions
\texttt{(add m n)} and \texttt{(add n m)} have been let-bound.

\begin{minipage}[t]{0.4\textwidth}
\begin{Verbatim}
let m = zero in
let n = (inc zero) in
(add (add (add m n) (add n m))
     (add (add n m) (add m n)))
\end{Verbatim}
\end{minipage}\hfill\begin{minipage}[t]{0.4\textwidth}
\begin{Verbatim}
let m = zero in
let n = (inc zero) in
let p = (add n m) in
let q = (add m n) in
(add (add q p) (add p q))
\end{Verbatim}
\end{minipage}
