\emph{Language Workbenches}, such as
Spoofax~\cite{DBLP:journals/software/WachsmuthKV14},
offer language designers an expressive environment in which to design,
implement, and deploy their \Acp{dsl}~\cite{hudak1996building}.
%
Principally speaking a language workbench~\cite{DBLP:conf/sle/ErdwegSVBBCGHKLKMPPSSSVVVWW13}
is a tool that supports:
description of a language's \emph{notation}---how we present a language's concrete syntax to users;
implementation of a language's \emph{semantics}---how we realise the language's behaviour;
and user interaction through an \emph{editor}.
%
Outside of these core criteria, various language workbenches
support language validation, testing, and composition.


Concurrently, the mechanised meta-theory research
programme~\cite{DBLP:conf/tphol/AydemirBFFPSVWWZ05,DBLP:journals/jfp/AbelAHPMSS19}
has seen a wealth of tools and techniques being developed
in the programming language theory community.
%
In particular, dependently typed languages such as
\Idris{}~\cite{DBLP:conf/ecoop/Brady21},
Agda~\cite{DBLP:conf/afp/Norell08},
and Coq~\cite{the_coq_development_team_2022_5846982}
have been widely used to formalise \acp{dsl}, and study their
meta-theoretical properties.
%
Efforts range from
studying specific core calculi~\cite{10.1145/3093333.3009866,DBLP:conf/cpp/RouvoetPKV20,DBLP:conf/mpc/ChapmanKNW19}
to building generic reasoning frameworks~\cite{DBLP:conf/cpp/StarkSK19,DBLP:journals/jfp/AllaisACMM21}.
%
These mechanised software verification projects, however, typically stop short
of actually building the frontend that would let users run these verified
language implementations.
If our verified language implementations type check, we might as well ship them too!
%
By becoming its own implementation language, \Idris{} has successfully
demonstrated that this is not an inescapable fate~\cite{DBLP:conf/ecoop/Brady21}.
%
But can we now claim that dependently typed languages qualify as
language workbenches?

\Velo{} is a minimal functional language that we have realised in \Idris{}
to showcase dependently typed techniques to implement and manipulate \acp{ir}.
%
This paper introduces \Velo{} but, most of all, seeks to show that
dependently typed languages make fine language workbenches.
%
We address both the core criteria and some optional extensions
highlighted by the language workbench challenge~\cite{DBLP:conf/sle/ErdwegSVBBCGHKLKMPPSSSVVVWW13} for what constitutes a language workbench.
%
Although not all of the optional criteria required for a language workbench is currently met by dependently typed languages,
we are convinced that with some additional engineering, taking advantage of existing work (e.g.\ Quickchick), more optional criteria can be satisfied.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:
