\subsection{Well-Typed Holes}
\label{sec:design:holes}

Holes are a special kind of placeholder that programmers can use for parts of the program they have not yet written.
%
In a typed language, each hole will be assigned a type based on the context it is used in.

\emph{Type-Driven Development}~\cite{DBLP:journals/pacmpl/OmarVCH19}
is a practice by which the user enters into a dialogue
\emph{with} the compiler to interactively build the program.
%
We can enable type-driven programming in part by providing special language support for holes and operations on them.
Such operations will include the ability to inspect, refine, compute with, and instantiate (with an adequately typed term) holes.
%
We believe that barebones language support for type-driven development
should at least include the ability to:
%
(1) inspect the type of a hole and the local context it appears in;
%
(2) instantiate a hole with an adequately typed term;
%
and as well
%
(3) safely evaluate programs that still contain holes.
%
\Velo{} provides all three.

\Idris{} elaborates holes as it encounters them by turning them into
global declarations with no associated definition.
%
Because of this design choice users cannot mention the same hole explicitly in different places to state their intention that these yet unwritten terms ought to be the same.
%
Users can refer to the hole's solution by its name,
but that hole is placed in one specific position
and it is from that position that \Idris{} infers its context.

In \Velo{}, however, we allow holes to be mentioned arbitrarily many times in
arbitrarily different local contexts.
%
In the following example, the hole \texttt{?h} occurs in two distinct contexts:
$\epsilon,\,a,\,x$ and $\epsilon,\,a,\,y$.

\begin{center}
  \holeexamplegraph{}
\end{center}

As a consequence, a term will only fit in that hole if it happens to live in the shared common prefix of these two contexts ($\epsilon,\,a$).
%
Indeed, references to $x$ will not make sense in $\epsilon,a,y$ and vice-versa for $y$.


Our elaborator proceeds in two steps.
%
First, a bottom-up pass records holes as they are found and, in nodes with multiple subterms, reconciles conflicting hole occurrences by computing the appropriate local context restrictions.
%
This process produces a list of holes, their types, and local contexts,
together with a \IdrisType{Holey} term that contains invariants ensuring
these collected holes do fit in the term.
%
Second, a top-down pass produces a core \IdrisType{Term} indexed by the list of \IdrisType{Meta} (a simple record type containing the hole's name, the context it lives in, and its type).
%
Hole occurences end up being assigned a thinning that embeds the metavariable's actual context into the context it appears in.
We discuss thinnings and their use in \Velo{} in Section~\ref{sec:compiler-pass}.

Although these intermediate representations are \Velo{}-specific, the technique
and invariants are general and can be reused by anyone wanting to implement
well-scoped holes in their functional \ac{dsl}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:
